---
title: "Activity Monitoring Data"
author: "Saideep Reddy"
date: "December 17, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


### Loading and preprocessing the data
```{r}
activitydata=read.csv("activity.csv",header=T)
activitydata$date=as.Date(activitydata$date)
```
### Mean total number of steps taken per day
```{r}
numofstep_perday=tapply(activitydata$steps,activitydata$date,sum)

```

```{r}
hist(numofstep_perday,breaks=25,main="Histogram of number of steps per day")

```

### Mean and median of the number of steps taken per day

```{r Mean and Median}
mean(numofstep_perday,na.rm=T)
median(numofstep_perday,na.rm=T)
```

### Average daily activity pattern

```{r}
avgsteps_perinterval=tapply(activitydata$steps,activitydata$interval,mean,na.rm=T)
plot(unique(activitydata$interval),avgsteps_perinterval,type="l",main="Avg steps per interval",xlab="interval",ylab="avg steps")
```

### Interval with maximum number of steps across all the days

```{r}
activitydata$interval[match(max(avgsteps_perinterval),avgsteps_perinterval)]
```

### Imputing missing values

Number of missing values in the dataset

```{r}
sum(is.na(activitydata))
```

Imputing the missing values with the average number of steps

```{r}

newactivitydata=activitydata
newactivitydata$steps[is.na(activitydata$steps)] = mean(activitydata$steps, na.rm=TRUE)

```

### Histogram of number of steps taken each day

```{r}
newnumofstep_perday=tapply(newactivitydata$steps,newactivitydata$date,sum)
hist(newnumofstep_perday,breaks=25,main="Histogram of number of steps per day")

```

### Mean and median

```{r}
mean(newnumofstep_perday)
median(newnumofstep_perday)

```

### Differences in activity patterns between weekdays and weekends

Factor to indicate weekend or not

```{r}
newactivitydata$weekday=weekdays(newactivitydata$date)
newactivitydata$weekend=(newactivitydata$weekday=="Saturday" | newactivitydata$weekday=="Sunday")

```

### Average number of steps taken during weekend and weekday

```{r, warning=FALSE}
library(data.table)
newactivitydata=data.table(newactivitydata)
newactivitydataweekday=newactivitydata[!(weekend)]
newactivitydataweekend=newactivitydata[(weekend)]

avgsteps_perinterval_weekday=tapply(newactivitydataweekday$steps,newactivitydataweekday$interval,mean)
plot(unique(activitydata$interval),avgsteps_perinterval_weekday,type="l",main="Weekday-Avg steps per interval",xlab="interval",ylab="avg steps")

avgsteps_perinterval_weekend=tapply(newactivitydataweekend$steps,newactivitydataweekend$interval,mean)
plot(unique(activitydata$interval),avgsteps_perinterval_weekend,type="l",main="Weekend-Avg steps per interval",xlab="interval",ylab="avg steps")


```